---
title: "Immune Paper - figures"
author: "Marc A.T. Teunis, PhD"
date: "`r Sys.time()`"
params:
  path_out: !r here::here('inst', 'final_figures_paper') 
  loose_images: !r here::here('inst', 'loose_images') 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{immune_viz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Packages
```{r}
library(tidyverse)
library(kinetics)
library(immunogram)
library(cowplot)
library(svglite)
library(gramlyr)
```

## Load data for analytes
```{r, panel_lines}
data_immune <- immunogram::data_immune
```

## Data checks
```{r}
## included subjects
levels(data_immune$subject)
## protocols in GRINTA!
levels(data_immune$protocol)
## analytes included in this analysis
unique(data_immune$analyte)
## time points measured
levels(data_immune$time)
```

## Missing values
```{r}
naniar::vis_miss(data_immune)
```

## Split data by analyte
```{r}
data_summary <- data_immune %>%
  group_by(protocol, time, analyte) %>%
  summarise(
    mean_conc = mean(concentration, na.rm = TRUE),
    sd = sd(concentration, na.rm = TRUE),
    n = n()) %>%
  arrange(protocol, analyte)
data_summary

data_summary$analyte <- as.factor(data_summary$analyte)
levels(data_summary$analyte)

data_split <- split(data_summary, 
                        droplevels(data_summary$analyte))

```

### Load analyte annotations
```{r}
data("analyte_annotations", package = "kinetics")
analyte_annotations
```

## Colours
```{r}
RColorBrewer::display.brewer.all()
palette <- RColorBrewer::brewer.pal(7, "Set1")
# GetColorHexAndDecimal <- function(color){
#   c <- col2rgb(color)
#   
#   sprintf("#%02X%02X%02X%3d%3d%3d", c[1],c[2],c[3],c[1],c[2],c[3])
# } 
# 
# black <- GetColorHexAndDecimal("black")
palette <- c("#000000", palette[c(1:3,4)])
```

## Line plots
```{r, all_lines, eval=TRUE}
image_directory <- params$loose_images
#dir.create(image_directory)

# argument for function test
analytes = data_immune$analyte %>% 
  unique()
immunogram::draw_lines(DF = data_split[[2]], palette_graph = palette)

map(
  .x = data_split,
  immunogram::draw_lines,
  palette_graph = palette
)


test_lines <- immuneGrinta::print_lines(analyte = analytes[1])
#test_lines
run_plotlist <- function(){
  plot_list <- map(
    .x = analytes, 
    .f = immunogram::print_lines)
  names(plot_list) <- levels(data_summary$analyte)
  names(plot_list)
}

```

## Panels
```{r, panels_plot_list, eval=FALSE}
## test
immunogram::print_lines_panel(analytes[4], x_pos = c(1,3))
```

## Prepare potlist for panels
```{r, plotlist}
## generate plotlist
analytes_figs <- c(
        
  "BASO_ABS",
  "LEUCO",
  "TROMBO",
  "EOS_ABS",
  "LYMFO_ABS_CORR",
  "MONO_ABS",
  "NEUTRO_ABS",
  "mcp1",
  "ip10",
  "tnf",
  "il6",
  "il10",
  "ifny",
  "gcsf",
  "mif"
  
)

analyte_data <- data_summary %>%
  dplyr::filter(analyte %in% analytes_figs) %>%
  droplevels()

levels(analyte_data$analyte)

panel_plot_list <- map(
  .x = levels(analyte_data$analyte), 
#  .y = x_pos_list,
  .f = immunogram::print_lines_panel,
  x_pos = c(1,3)
  )

# panel_plot_list
## check x_pos for GLU_NS
names(panel_plot_list) <- levels(analyte_data$analyte)
names(panel_plot_list)

## save to disk for use in the package
# usethis::use_data(panel_plot_list)

```

## Final panels

### Figure 1
```{r}


## names(panel_plot_list) <- analytes_figs

p1_a <- panel_plot_list[["LYMFO_ABS_CORR"]] 
p1_b <- panel_plot_list[["MONO_ABS"]]
p1_c <- panel_plot_list[["NEUTRO_ABS"]]
p1_d <- panel_plot_list[["EOS_ABS"]]
p1_e <- panel_plot_list[["BASO_ABS"]]
p1_f <- panel_plot_list[["TROMBO"]]

## dimensions
height <- 28
width <- 16
resolution <- 600

# https://cran.r-project.org/web/packages/cowplot/vignettes/shared_legends.html

horizontal_label_offset <- -7
vertical_label_offset <- 4
label_size <- 20

figure_1 <- cowplot::plot_grid(p1_a + theme(legend.position="none"),
           p1_b + theme(legend.position="none"),
           p1_c + theme(legend.position="none"),
           p1_d + theme(legend.position="none"),
           p1_e + theme(legend.position="none"),
           p1_f + theme(legend.position="none"),
           align = 'vh',
           labels = c("A", "B", "C", "D", "E", "F"),
           hjust = horizontal_label_offset, 
           vjust = vertical_label_offset,
           ncol = 2,
           label_size = label_size
           )

#cfigure_1


legend_1 <- get_legend(p1_a + theme(legend.position="bottom"))

p_1 <- plot_grid(legend_1, figure_1, nrow = 2, rel_heights = c(0.05, 1))
p_1


##kinetics::save_in_image_directory(filename = "test2.svg", height = 26, width = 7)
ggsave(p_1, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_1.svg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)

ggsave(p_1, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_1.png", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)


ggsave(p_1, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_1.jpeg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)


```

### Figure 2

```{r}

```


### Figure 3
a) MCP1
b) Ip10
c) TNf
```{r}
## names(panel_plot_list) <- analytes_figs

p3_a <- panel_plot_list[["mcp1"]] + 
  ggtitle("MCP-1") +
  ylab("Conc. (pg/ml)")
p3_b <- panel_plot_list[["ip10"]] + 
  ggtitle("IP-10") +
  ylab("Conc. (pg/ml)")
p3_c <- panel_plot_list[["tnf"]] + 
  ggtitle("TNF-alpha") +
  ylab("Conc. (pg/ml)")

## dimensions
height <- 10
width <- 24
resolution <- 600

# https://cran.r-project.org/web/packages/cowplot/vignettes/shared_legends.html

horizontal_label_offset <- -7
vertical_label_offset <- 4
label_size <- 20

figure_3 <- cowplot::plot_grid(p3_a + theme(legend.position="none"),
           p3_b + theme(legend.position="none"),
           p3_c + theme(legend.position="none"),
           align = 'vh',
           labels = c("A", "B", "C"),
           hjust = horizontal_label_offset, 
           vjust = vertical_label_offset,
           nrow = 1,
           label_size = label_size
           )

#cfigure_1


legend_3 <- get_legend(p3_a + theme(legend.position="top"))

p_3 <- plot_grid(legend_3, figure_3, nrow = 2, 
                 rel_heights = c(0.1, 1)
                # rel_widths = c(0.02, 1.4)
                 )
p_3


##kinetics::save_in_image_directory(filename = "test2.svg", height = 26, width = 7)
ggsave(p_3, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_3.svg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)

ggsave(p_3, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_3.png", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)


ggsave(p_3, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_3.jpeg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)

```



### Figure 4
a) IL-6
b) IL-10
c) IFN-g
d) G-CSF
e) MIF

```{r}
## names(panel_plot_list) <- analytes_figs
p4_a <- panel_plot_list[["il6"]] + 
  ggtitle("IL-6") +
  ylab("Conc. (pg/ml)")
p4_b <- panel_plot_list[["il10" ]] + 
  ggtitle("IL-10") +
  ylab("Conc. (pg/ml)")
p4_c <- panel_plot_list[["ifny"]] + 
  ggtitle("IFN-gamma") +
  ylab("Conc. (pg/ml)")
p4_d <- panel_plot_list[["gcsf"]] + 
  ggtitle("G-CSF") +
  ylab("Conc. (pg/ml)")
p4_e <- panel_plot_list[["mif"]] + 
  ggtitle("MIF") +
  ylab("Conc. (pg/ml)")

## dimensions
height <- 20
width <- 24
resolution <- 600

# https://cran.r-project.org/web/packages/cowplot/vignettes/shared_legends.html

horizontal_label_offset <- -7
vertical_label_offset <- 4
label_size <- 20

figure_4 <- cowplot::plot_grid(p4_a + theme(legend.position="none"),
           p4_b + theme(legend.position="none"),
           p4_c + theme(legend.position="none"),
           p4_d + theme(legend.position="none"),
           p4_e + theme(legend.position="none"),
           align = 'vh',
           labels = c("A", "B", "C", "D", "E"),
           hjust = horizontal_label_offset, 
           vjust = vertical_label_offset,
           nrow = 2,
           label_size = label_size
           )

#cfigure_1


legend_4 <- get_legend(p4_a + theme(legend.position="top"))

p_4 <- plot_grid(legend_4, figure_4, nrow = 2, 
                 rel_heights = c(0.05, 1)
                # rel_widths = c(0.02, 1.4)
                 )
p_4


##kinetics::save_in_image_directory(filename = "test2.svg", height = 26, width = 7)
ggsave(p_4, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_4.svg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)

ggsave(p_4, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_4.png", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)


ggsave(p_4, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "fig_4.jpeg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)

```

## NK cells

### Packages
```{r}
library(tidyverse)
library(gramlyr)
library(toolboxr)
library(readxl)
library(naniar)
library(immunogram)
```

### Data
```{r}
data_nk <- immunogram::data_nk
data_nk
```

## Inspect
```{r}
map(data_nk, typeof)
names(data_nk)
sum(is.na(data_nk))

```

## Data Clean

Observations for ID2 and protocol 4 are invalid!!!

 - used , instead of .
 - % in value

```{r}
## data_nk
data_nk_clean <- data_nk %>%
  dplyr::filter(Subject != 2, Protocol != 4)

data_nk_clean[, -c(1:9, 29)] <- map(data_nk_clean[, -c(1:9, 29)],
                                as.numeric)
names(data_nk_clean) <- names(data_nk_clean) %>%
  tolower() %>%
  str_replace_all(pattern = "/", replacement = "_")

data_names <- names(data_nk_clean) %>% enframe()

data_nk_clean <- data_nk_clean %>%
  mutate(time_hrs = time/60)
data_nk_clean$time_hrs <- as_factor(data_nk_clean$time_hrs)
levels(data_nk_clean$time_hrs)
```

## Missing 
```{r}
naniar::vis_miss(data_nk)
sum(is.na(data_nk_clean))
naniar::vis_miss(data_nk_clean)
```

# NK Cells

## Colours
```{r}
RColorBrewer::display.brewer.all()
palette <- RColorBrewer::brewer.pal(7, "Set1")
palette <- c("#000000", palette[c(1:3,4)])
```

## Total percentage NK - all data
```{r}
data_nk_clean %>%
  dplyr::select(perctotalnk, time, subject, protocol) %>%
  ggplot(aes(x = subject,
             y = perctotalnk)) +
  geom_point(aes(colour = protocol)) +
  facet_wrap(~time)

```

## CD57d/CD16

d = dim -> inflammation
b = bright -> cytokine producing NK

## Bright / Dim separated

### Function to plot
```{r}
#' @title Function that plots specific flow cytometric subsets for NK cells




plot_population_percentage <- function(data, analyte){
  
plot <- data %>%
  group_by(protocol, time_hrs) %>%
  summarise(mean_total_pop = mean(get(analyte), na.rm = TRUE)) %>%
  ggplot(aes(x = time_hrs,
             y = mean_total_pop)) +
  geom_line(aes(group = protocol, colour = protocol), size = 2) +
  geom_point(data = data_nk_clean, 
             aes(x = time_hrs, 
             y = get(analyte),
             colour = protocol), position = "jitter") 
  return(plot)
}

```

### Inspect
```{r}
data_nk_clean %>%
  dplyr::select(
    subject,
    protocol,
    time,
    perccd16d_cd56bnk,
    perccd16b_cd56bnk,
    perccd16b_cd56dnk
  ) %>%
  pivot_longer(
    cols = c(
      perccd16d_cd56bnk,
      perccd16b_cd56bnk, 
      perccd16b_cd56dnk),
    values_to = "percentage",
    names_to = "population") %>%
  group_by(protocol, time, population) %>%
  summarise(mean_perc = mean(percentage, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = mean_perc)) +
  geom_point(aes(colour = population)) +
  geom_line(aes(group = population, colour = population)) +
  facet_wrap(~protocol)
```

```{r}

plot_labels <- c(
  `a` = "A",
  `b` = "B",
  `c` = "C"
)

nk_panel <- data_nk_clean %>%
  dplyr::select(
    subject,
    protocol,
    time,
    perccd16d_cd56bnk,
    perccd16b_cd56bnk,
    perccd16b_cd56dnk
  ) %>%
  pivot_longer(
    cols = c(
      perccd16d_cd56bnk,
      perccd16b_cd56bnk, 
      perccd16b_cd56dnk),
    values_to = "percentage",
    names_to = "population") %>%
  mutate(time_hrs = time/60) %>%
  mutate(population_new = fct_recode(as_factor(population), 
      "CD16dim/CD56bright NK" = "perccd16d_cd56bnk",
      "CD16bright/CD56bright NK" = "perccd16b_cd56bnk", 
      "CD16bright/CD56dim NK" = "perccd16b_cd56dnk")) %>%
  group_by(protocol, time_hrs, population_new) %>%
  summarise(mean_perc = mean(percentage, na.rm = TRUE)) %>%
  ggplot(aes(x = time_hrs, y = mean_perc)) +
#  geom_point(aes(colour = protocol)) +
  geom_line(aes(group = protocol, colour = protocol), size = 2) +
  ggplot2::geom_point(aes(shape = protocol, colour = protocol),
                        size = 3,
                        position = position_dodge(width = 0.2),
                        show.legend = FALSE) +
    ggplot2::scale_colour_manual(labels = c("Rest",
                                            "70% Wmax",
                                            "70% Wmax/DH",
                                            "50% Wmax",
                                            "55%/85% Wmax"),
                                 values = palette)   +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed") +
    ggplot2::geom_vline(xintercept = 1, linetype = "dashed") +
    geom_label(aes(x = 0, y = 0,
                   label = "T0"), colour = "black") +
    geom_label(aes(x = 1, y = 0,
                   label = "T1"), colour = "black") +
    ggplot2::scale_shape_manual(values=c(5,6, 15:17)) +
    kinetics::theme_individual() +
    ggplot2::xlab("Time (hours)") +
    ggplot2::ylab("Mean % of total") +
    facet_wrap(~ population_new)

ggsave(
  nk_panel,
  filename = file.path(params$path_out, "nk_cells_panel.png"),
  width = 50, height = 14, dpi = 600, units = "cm"
  )

ggsave(
  nk_panel,
  filename = file.path(params$path_out, "nk_cells_panel.svg"),
  width = 50, height = 14, dpi = 600, units = "cm"
  )


ggsave(
  nk_panel,
  filename = file.path(params$path_out, "nk_cells_panel.jpeg"),
  width = 50, height = 14, dpi = 600, units = "cm"
  )

```

```{r}
data_names

percentages_list <- list(
  data_names[13,2] %>% as.character(),
  data_names[17,2] %>% as.character(),
  data_names[21,2] %>% as.character()
)

percentages_list

map(
  percentages_list,
  data = data_nk_clean,
  plot_population_percentage) -> plot_nk_list
```

### Figure 2A
```{r}
p_2a <- plot_nk_list[[1]] +
    ggplot2::geom_point(aes(shape = protocol),
                        size = 3,
                        position = position_dodge(width = 0.2),
                        show.legend = FALSE) +
    ggplot2::scale_colour_manual(labels = c("Rest",
                                            "70% Wmax",
                                            "70% Wmax/DH",
                                            "50% Wmax",
                                            "55%/85% Wmax"),
                                 values = palette)   +
    ggplot2::geom_vline(xintercept = 1, linetype = "dashed") +
    ggplot2::geom_vline(xintercept = 2, linetype = "dashed") +
    geom_label(aes(x = 1, y = 0,
                   label = "T0"), colour = "black") +
    geom_label(aes(x = 2, y = 0,
                   label = "T1"), colour = "black") +
    ggplot2::scale_shape_manual(values=c(5,6, 15:17)) +
    kinetics::theme_individual() +
    ggplot2::xlab("Time (hours)") +
    ggplot2::ylab("Mean % of total") +
    ggtitle("CD16dim/CD56bright NK")

p_2a

ggsave(
  filename = file.path(
    params$path_out, "2a_cd16d_cd56b_nk.png"),
    plot = p_2a, dpi = 300, width = 28, height = 16, units = "cm"
  ) 
```

Figure 2B
```{r}
p_2b <- plot_nk_list[[2]] +
    ggplot2::geom_point(aes(shape = protocol),
                        size = 3,
                        position = position_dodge(width = 0.2),
                        show.legend = FALSE) +
    ggplot2::scale_colour_manual(labels = c("Rest",
                                            "70% Wmax",
                                            "70% Wmax/DH",
                                            "50% Wmax",
                                            "55%/85% Wmax"),
                                 values = palette)   +
    ggplot2::geom_vline(xintercept = 1, linetype = "dashed") +
    ggplot2::geom_vline(xintercept = 2, linetype = "dashed") +
    geom_label(aes(x = 1, y = 0,
                   label = "T0"), colour = "black") +
    geom_label(aes(x = 2, y = 0,
                   label = "T1"), colour = "black") +
    ggplot2::scale_shape_manual(values=c(5,6, 15:17)) +
    kinetics::theme_individual() +
    ggplot2::xlab("Time (hours)") +
    ggplot2::ylab("Mean % of total") +
    ggtitle("CD16bright/CD56bright NK")
   
p_2b

ggsave(
  filename = file.path(
    params$path_out, "2b_cd16b_cd56b_nk.png"),
    plot = p_2b, dpi = 300, width = 28, height = 16, units = "cm"
  ) 
```

### Figure 2C
```{r}
p_2c <- plot_nk_list[[1]] +
    ggplot2::geom_point(aes(shape = protocol),
                        size = 3,
                        position = position_dodge(width = 0.2),
                        show.legend = FALSE) +
    ggplot2::scale_colour_manual(labels = c("Rest",
                                            "70% Wmax",
                                            "70% Wmax/DH",
                                            "50% Wmax",
                                            "55%/85% Wmax"),
                                 values = palette)   +
    ggplot2::geom_vline(xintercept = 1, linetype = "dashed") +
    ggplot2::geom_vline(xintercept = 2, linetype = "dashed") +
    geom_label(aes(x = 1, y = 0,
                   label = "T0"), colour = "black") +

    geom_label(aes(x = 2, y = 0,
                   label = "T1"), colour = "black") +
    ggplot2::scale_shape_manual(values=c(5,6, 15:17)) +
    kinetics::theme_individual() +
    ggplot2::xlab("Time (hours)") +
    ggplot2::ylab("Mean % of total") +
    ggtitle("CD16bright/CD56dim NK")
   
p_2c

ggsave(
  filename = file.path(
    params$path_out, 
    "2c_cd16b_cd56d_nk.png"),
    plot = p_2c, 
    dpi = 300, 
    width = 28, 
    height = 16, 
    units = "cm"
  ) 

```
